Holomorphic field theories and higher algebra

### 2024-09-20 draft.tex
rearranging subsections in 2.2 and adding subsections in 6

\epigraph{The mathematical language of classical physics is based upon real numbers \dots The mathematical language of quantum physics is based upon complex numbers, and it would be natural to expect that the complex analytic and the algebraic geometry should replace the differential geometry of the classical period.}{Yuri Manin from the introduction of {\it Frobenius Manifolds, Quantum Cohomology, and Moduli Spaces}}

@@

\begin{eg}[Holomorphic Chern-Simons theory]
\label{eg: holCS}
\owen{This feels rather convoluted. Can we make it simpler?}
Pick a {\em complex} Lie group $G$, 
and equip its Lie algebra $\fg$ with a complex-linear, $G$-invariant nondegenerate symmetric bilinear form~$\kappa$.
Let $M$ be a three-dimensional \textit{complex} manifold
equipped with a holomorphic volume form $\Omega \in \Omega^{3,0}(M)$; 
in other words, $M$ is a Calabi--Yau manifold.
(This condition is analogous to being an oriented 3-manifold.)
Given a smooth principal $G$-bundle $P \to M$,
there is a moduli space of holomorphic structures on~$P$.
Suppose there exists at least one holomorphic structure $\cP \to M$,
which has an associated Dolbeault operator (or $\dbar$-connection) $\dbar_{\cP}$.
Any other holomorphic bundle structure determines a Dolbeault operator,
but our reference point $\dbar_{\cP}$ provides an isomorphism
\[
{\rm Conn}^\dbar_G(P) \cong \Omega^{0,1}(M, {\rm ad}(P)),
\]
i.e., we express another $\dbar$-connection as $\dbar_{\cP} + A$ where $A$ is a $(0,1)$-form valued in the adjoint bundle.
The equation of motion is again a Maurer--Cartan equation
\beqn
\dbar_{\cP} A + \frac12 [A,A] = 0 ,
\eeqn
or equivalently the $(0,2)$-term of the curvature vanishes.
This PDE is the Euler--Lagrange equation for the action
\beqn
S_{hCS} (A) = \int_X \Omega \wedge CS(A) 
\eeqn
where $CS(A)$ is exactly as in~\eqref{eqn:CS}.\footnote{Notice that due to the presence of the holomorphic volume form $\Omega$, only the $\dbar A$ component of $\d A$ contributes to the integrand of the action functional, because $\Omega \wedge \partial A$ vanishes.} 
As in the case of Chern-Simons theory, one must quotient by the appropriate group of gauge transformations to obtain the correct moduli space, 
which is here the moduli stack of holomorphic principal $G$-bundles whose underlying smooth structure is given by $P \to M$.
\end{eg}

@@

@@

\subsubsection{Holomorphic field theories as chiral sectors}

\owen{I'm a little lost about the motivation for this section, 
given the earlier scalar field section. This C theory looks interest so maybe we can figure out a natural place to put it?}

A typical scalar field theory is the free boson which exists on any Riemannian manifold $\Sigma$.
Its action functional is simply $\int_\Sigma \phi \triangle \phi$ where $\phi$ is a smooth function on $\Sigma$.
We saved this example for last since its holomorphic analog is a bit more elusive than the previous examples.

We start with complex dimension~$1$.
The usual scalar field theory can be written as 
\[
\int_{\Sigma} \phi \wedge \del \dbar \phi, 
\]
where $\Sigma$ is a Riemann surface.
This theory is not holomorphic, but we can force it to be by demanding the constraint $\dbar \phi = 0$.
Importantly, this constraint is \textit{not} imposed by an equation of motion.
When we impose this constract, we can rewrite the action functional in terms of the $(1,0)$-form $\alpha \in \Omega^{1,0}(\Sigma)$ as
\beqn\label{eqn:scalar1}
\int_\Sigma \alpha \wedge \dbar \del^{-1} \alpha ,
\eeqn
where $\alpha$ plays the role of $\partial \phi$.
\owen{Am I interpreting that correctly?} 
This is formally similar to the kinetic term in the action functional of Kodaira--Spencer theory \eqref{eqn:KS1}.

There is a higher-dimensional analog of the chiral sector of the scalar field.
Suppose that $X$ is a complex manifold of \textit{odd} dimension $2n+1$.
The fields are the $(n+1,n)$-forms $\alpha \in \Omega^{n+1,n}(X)$ that satisfy the constraint
\beqn
\del \alpha = 0.
\eeqn
Note that when $n=0$, so $X$ is simply a Riemann surface~$\Sigma$, this constraint is automatic.
The action functional is the same as~\eqref{eqn:scalar1}.
When $n=1$ this functional is identical to the kinetic part of the Kodaira--Spencer action.\footnote{Indeed, when $X$ is a Calabi--Yau three-fold there is an isomorphism $\Omega^{2,1}(X) \cong \Omega^{0,1}(X, \T_X)$.}
Generally, this theory is the holomorphic part of a higher-form scalar theory, with action
\beqn
\int_X \|C\|^2 \op{dvol} = \int_X C \wedge \d \star \d C , 
\eeqn
where $C \in \Omega^{2n+1}(X)$.
For more details we refer to~\cite{GRWwzw}.

@@

@@

@@

>>>>

\epigraph{The mathematical language of classical physics is based upon real numbers \dots. The mathematical language of quantum physics is based upon complex numbers, and it would be natural to expect that the complex analytic and the algebraic geometry should replace the differential geometry of the classical period.}{Yuri Manin from the introduction of {\it Frobenius Manifolds, Quantum Cohomology, and Moduli Spaces}}

@@

\begin{eg}[Holomorphic Chern-Simons theory]
\label{eg: holCS}
\owen{This feels rather convoluted. Can we make it simpler?}
Pick a {\em complex} Lie group $G$, 
and equip its Lie algebra $\fg$ with a complex-linear, $G$-invariant nondegenerate symmetric bilinear form~$\kappa$.
Let $M$ be a three-dimensional \textit{complex} manifold
equipped with a holomorphic volume form $\Omega \in \Omega^{3,0}(M)$; 
in other words, $M$ is a Calabi--Yau manifold.
(This condition is analogous to being an oriented 3-manifold.)
Given a smooth principal $G$-bundle $P \to M$,
there is a moduli space of holomorphic structures on~$P$.
Suppose there exists at least one holomorphic structure $\cP \to M$,
which has an associated Dolbeault operator (or $\dbar$-connection) $\dbar_{\cP}$.
Any other holomorphic bundle structure determines a Dolbeault operator,
but our reference point $\dbar_{\cP}$ provides an isomorphism
\[
{\rm Conn}^\dbar_G(P) \cong \Omega^{0,1}(M, \fg_P),
\]
i.e., we express another $\dbar$-connection as $\dbar_{\cP} + A$ where $A$ is a $(0,1)$-form valued in the adjoint bundle.
The equation of motion is again a Maurer--Cartan equation
\beqn
\dbar_{\cP} A + \frac12 [A,A] = 0 ,
\eeqn
or equivalently the $(0,2)$-term of the curvature vanishes.
This PDE is the Euler--Lagrange equation for the action
\beqn
S_{hCS} (A) = \int_M \Omega \wedge CS(A) 
\eeqn
where $CS(A)$ is exactly as in~\eqref{eqn:CS}.\footnote{Notice that due to the presence of the holomorphic volume form $\Omega$, only the $\dbar A$ component of $\d A$ contributes to the integrand of the action functional, because $\Omega \wedge \partial A$ vanishes.} 
As in the case of Chern-Simons theory, one must quotient by the appropriate group of gauge transformations to obtain the correct moduli space, 
which is here the moduli stack of holomorphic principal $G$-bundles whose underlying smooth structure is given by $P \to M$.
\end{eg}

@@

\subsubsection{Interlude on holomorphic field theories as chiral sectors}

Every holomorphic theory we have introduced so far has a ``kinetic term'' (the quadratic term in the action functional) that only involves $\dbar$.
Thus the linearized equations of motion always have the form $\dbar \phi = 0$,
so that they manifestly relate to holomorphic functions (or sections).
There are, however, holomorphic theories with a different flavor.

\begin{comment}
A typical scalar field theory is the free boson which exists on any Riemannian manifold $\Sigma$.
Its action functional is simply $\int_\Sigma \phi \triangle \phi$ where $\phi$ is a smooth function on $\Sigma$.
We saved this example for last since its holomorphic analog is a bit more elusive than the previous examples.
\end{comment}

We start with complex dimension~$1$.
The usual scalar field theory can be written as 
\[
\int_{\Sigma} \phi \wedge \del \dbar \phi, 
\]
where $\Sigma$ is a Riemann surface.
Using integration by parts, one can use instead the action functional
\[
\int_{\Sigma} \partial \phi \wedge \dbar \phi.
\]
This theory is not holomorphic, but we can force it to be holomorphic by demanding that $\dbar \phi = 0$.
Importantly, this constraint is \textit{not} an equation of motion.
When we impose this constraint, we can rewrite the action functional in terms of the $(1,0)$-form $\alpha \in \Omega^{1,0}(\Sigma)$ as
\beqn\label{eqn:scalar1}
\int_\Sigma \alpha \wedge \dbar (\del^{-1} \alpha) ,
\eeqn
where $\alpha$ plays the role of $\partial \phi$.
In this theory the kinetic term is the whole action,
and it has a rather different flavor from the examples we have seen so far.\footnote{This action functional is a precursor to the ``holomorphic gravity'' theory we meet next, 
known as Kodaira--Spencer theory.}
One might call this the {\em chiral sector} of the free scalar theory,
as it consists of the complex-valued scalar fields satisfying a ``chiral'' constraint.

There is a higher-dimensional analog of the chiral sector of the scalar field.
Let $X$ be a complex manifold of dimension $2n+1$, i.e., odd.
The fields are the $(n+1,n)$-forms $\alpha \in \Omega^{n+1,n}(X)$ that satisfy the constraint
\beqn
\del \alpha = 0.
\eeqn
(When $n=0$ and $X$ is simply a Riemann surface, 
this constraint is automatic.)
The action functional is the same as~\eqref{eqn:scalar1}.
Generally, this theory is the holomorphic part of a higher-form scalar theory, with action
\beqn
\int_X \|C\|^2 \op{dvol} = \int_X \d C \wedge \star \d C , 
\eeqn
where $C \in \Omega^{2n+1}(X)$.
For more details we refer to~\cite{GRWwzw}.

@@

@@

\owen{In this section, or maybe the next, we should mention the recent papers by physicists (notably Gaiotto) about higher bracket structures, and we should indicate how to connect with our approach here.}

@@

\subsection{No divergences}

@@

\subsection{Anomalies that obstruct quantization}
